---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Zoning for Autonomous Vehicles

<!-- badges: start -->
<!-- badges: end -->

The goal of zav (Zoning for Autonomous Vehicles) is to show the implementation of the problem instance generation and simulation experiment conducted for the project.

## Installation

You can install the development version of zav from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Rosenkrands/zav")
```

## Example

This is a basic example which shows how to generate  a problem instance:

```{r example, fig.dim=c(4,4)}
library(zav)

instance <- generate_2d_instance()
plot_point(instance = instance)
```

Here it is shown how the instance can be solved by the KMeans solution approach:

```{r solution}
solution <- solve_kmeans(instance, no_of_centers = 5)
plot_bases(instance = instance, solution = solution)
```

