---
output: github_document
editor_options: 
  markdown: 
    wrap: sentence
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300
)
```

# Zoning for Autonomous Vehicles

<!-- badges: start -->

<!-- badges: end -->

The goal of zav (Zoning for Autonomous Vehicles) is to show the implementation of the problem instance generation and simulation experiment conducted for the project.

## Installation

You can install the development version of zav from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# note that there is a GitHub only dependency
# devtools::install_github("cuhklinlab/SWKM")
devtools::install_github("Rosenkrands/zav")
```

## Instance generation

To generate a problem instance we can utilize the `generate_2d_instance` function.

```{r example}
library(zav)
library(ggplot2)
library(dplyr)

instance <- generate_2d_instance(no_of_points = 100)

plot_point(instance)
```

## Zoning solutions

### Weighted K-Means

The below code chunk shows how we can utilize the `solve_wkmeans` function to generate a solution for our problem instance.

```{r solution}
solution_wkm <- solve_wkmeans(instance, no_of_centers = 5, type = "swkm")
plot_bases(solution = solution_wkm)
```

### Genetic Algorithm

The below chunk shows how we can utilize the `solve_ga` function to generate a solution for our problem instance.

First we need to precalculate centroids to use as input for the GA.

```{r, cache=T}
centroids <- grid_centroids(instance, dimension = 5)
```

Having now the centroids and distances between demand points and centroids, we are able to give this as input for the GA.
As a default the GA will have a maximum of 10 iterations for demonstration purposes, in reality we would have a much higher number of iterations.

```{r, cache=T}
solution <- solve_ga(instance, centroids, no_of_centers = 5, obj = "SAFE")
```

```{r}
plot_bases(solution)
```

## Simulation

Here it is shown how you can make a simulation based on the solution just found:

```{r simulation, cache = T}
simulation_result <- simulation(
  solution = solution_wkm,
  seed = 1,
  n_replications = 1,
  flight = "zoned",
  queue = T,
  max_dist = 1000000,
  LOS = 600,
  warmup = 0,
  speed_agent = .25,
  verbose = F
)
```

```{r distances, fig.dim=c(8,3)*1.1}
simulation_result$metrics[[1]]$distances %>%
  ggplot(aes(x = time, 
             y = distance,
             color = paste0(id1,'-',id2))) +
  geom_line() +
  labs(color = "Pair") +
  theme_bw() +
  theme(legend.position = "none")
```
