---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 300
)
```

# Zoning for Autonomous Vehicles

<!-- badges: start -->
<!-- badges: end -->

The goal of zav (Zoning for Autonomous Vehicles) is to show the implementation of the problem instance generation and simulation experiment conducted for the project.

## Installation

You can install the development version of zav from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
# note that there is a GitHub only dependency
# devtools::install_github("cuhklinlab/SWKM")
devtools::install_github("Rosenkrands/zav")
```

## Instance generation

To generate a problem instance we can utilize the `generate_2d_instance` function.

```{r example}
library(zav)
library(ggplot2)

instance <- generate_2d_instance(no_of_points = 100)
# plot_point(instance = instance)
ggplot2::ggplot(instance$data) +
    ggplot2::geom_point(ggplot2::aes(x,y,size=`Arrival rate`),
                        shape = 21, fill = alpha("black", .2)) +
    ggplot2::theme_void()
```

## Zoning solutions

### Weighted K-Means

The below code chunk shows how we can utilize the `solve_wkmeans` function to generate a solution for our problem instance.

```{r solution}
solution_wkm <- solve_wkmeans(instance, no_of_centers = 5, type = "swkm")
plot_bases(solution = solution_wkm)
```

### Genetic Algorithm

The below chunk shows how we can utilize the `solve_ga` function to generate a solution for our problem instance.

First we need to precalculate centroids to use as input for the GA.

```{r, cache=T}
centroids <- grid_centroids(instance, dimension = 5)
```

Having now the centroids and distances between demand points and centroids, we are able to give this as input for the GA.
As a default the GA will have a maximum of 10 iterations for demonstration purposes, in reality we would have a much higher number of iterations. 

```{r, cache=T}
solution <- solve_ga(instance, centroids, no_of_centers = 5, obj = "SAFE")
```

```{r}
plot_bases(solution)
```

## Simulation

Here it is shown how you can make a simulation based on the solution just found:

```{r simulation, cache = T}
#' The below line is for testing purposes
#' solution = solution; seed = 1;n_replications = 1;flight = "zoned";max_dist = 1000000;LOS = 600;warmup = 0;speed_agent = .25;verbose = F
simulation_result <- simulation(
  solution = solution_wkm,
  seed = 1,
  n_replications = 1,
  flight = "zoned",
  queue = T,
  max_dist = 1000000,
  LOS = 600,
  warmup = 0,
  speed_agent = .25,
  verbose = F
)
```

```{r}
# simulation_result$metrics[[1]]$response_time_performance
```


```{r distances}
simulation_result$metrics[[1]]$distances %>%
  ggplot(aes(x = time, 
             y = distance,
             color = paste0(id1,'-',id2))) +
  geom_line() +
  labs(color = "Pair") +
  theme_bw() +
  theme(legend.position = "none")
```

## Experiment Results

To get analyze the results from the experiment conducted for the project we do the following.

```{r}
results <- experiment_results()
```

The `results` variable is a list of 3 `tibbles` containing information about instances, solutions and simulations respectively.

### Instances

If we take a look at the instances first we can see that there is a total of `r nrow(results$instance)` instances.
These are distributed across the arrival rate distributions and arrival rate variances as seen below.

```{r, message = F}
results$instance |> 
  dplyr::group_by(`Arrival rate distribution`, `Arrival rate variance`) |> 
  dplyr::summarise(n = dplyr::n())
```

In the illustration below we can see the difference in arrival rate distribution between the uniform and unbalanced distribution.

```{r, echo = F, fig.height=3}
uniform_instance <- readRDS(system.file("extdata", "instances/010E.rds", package = "zav"))
reciprocal_instance_paper <- readRDS(system.file("extdata", "instances/01HQ.rds", package = "zav"))

cowplot::plot_grid(
    zav::plot_point(uniform_instance) + ggplot2::ggtitle("Uniform distribution"),
    zav::plot_point(reciprocal_instance_paper) + ggplot2::ggtitle("Unbalanced distribution")
)
```

### Solutions

Taking a look at the solutions we see that there are a total of `r nrow(results$solution)` solutions.
These are distributed across:

* `Solution method` that have `r nrow(unique(results$solution[, "Solution method"]))` levels: `r unique(results$solution[, "Solution method"][[1]])`.
* `Number of UAVs` that have `r nrow(unique(results$solution[, "Number of UAVs"]))` levels: `r unique(results$solution[, "Number of UAVs"][[1]])`.

#### Solution comparsion

If we start by comparing ga-tot and wkm-swkm we see that the latter is able to outperform the first, with the high number of UAVs.

```{r, echo=F, message=F}
results$solution |> 
  dplyr::filter(`Solution method` != "ga-safe") |> 
  dplyr::group_by(`Solution method`, `Arrival rate variance`, `Arrival rate distribution`, `Number of UAVs`) |> 
  dplyr::summarise(sd = sd(TOT), TOT = mean(TOT)) |> 
  ggplot2::ggplot(ggplot2::aes(x = `Number of UAVs`, y = TOT, group = `Solution method`, color = `Solution method`)) +
  ggplot2::geom_point() +
  ggplot2::geom_line() +
  ggplot2::geom_errorbar(ggplot2::aes(ymin = TOT - sd, ymax = TOT + sd), width = .05) +
  ggplot2::facet_grid(`Arrival rate distribution`~`Arrival rate variance`, scales="free_y", labeller = ggplot2::label_both)
```

Next we will check the similarity of total arrival rate for uniform and unbalanced instances.

```{r}
results$instance |> 
  dplyr::group_by(`Arrival rate distribution`) |> 
  dplyr::summarise(mean(`Total arrival rate`))
```

The total arrival rate is not at all the same, so should maybe be adjusted after looking at the simulation results!

### Simulations

work in progress...
